A community-maintained standard library of population genetic simulations

### 2019-12-19 popSimManu.tex

\section*{Results}
\paragraph{The \stdpopsim library}
The central contribution of the PopSim consortium currently is \stdpopsim, a
community-maintained library of empirical genome data and population genetics simulation
models, providing easy access to realistic simulations. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans, \emph{P.~abelii},
\emph{D.~melanogaster}, \emph{A.~thaliana}, and \emph{E.~coli}.
% RNG: Abbreviate consistently with introduction
A species definition consists of two key elements.  Firstly, the library defines
some basic information about the species' genome, including information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as recombination maps, which model empirically-observed
heterogeneity along chromosomes. As such maps are often large and hard
to find, we provide a mechanism to automatically download (and cache locally) a
set of previously published maps (Fig. \ref{fig:cartoon}B). For example, in humans we support the
HapMapII~\citep{international2007second} and
deCODE~\citep{kong2010fine} genetic maps. The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios that have influenced present-day patterns of
genetic variation.

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \stdpopsim has a Python API and a user-friendly
command line interface (Fig.~\ref{fig:cartoon}C), allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the “tree sequence”
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed efficiently using the \tskit library
\citep{kelleher2016efficient,kelleher2018efficient}. Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine. We plan to include other simulation
engines to support scenarios that cannot be modelled under the coalescent framework,
and have already implemented \texttt{SLiM} \citep{haller2019slim} as
an alternative back-end.

The \stdpopsim command line interface, by default, outputs citation information
for the models, genetic maps and simulation engines used in any particular run.
We hope that this will encourage users to appropriately acknowledge the
resources used in published work and authors
of simulation models to contribute to our ongoing community-driven development process.
Together with the \stdpopsim version number, this citation information will result in a reasonably well-documented and reproducible simulation workflow.

>>>>

\section*{Results}
\paragraph{The \stdpopsim library}
The first contribution of the PopSim consortium is \stdpopsim, a
community-maintained library of empirical genome data and population genetics simulation
models. Figure \ref{fig:cartoon} shows a graphical
representation of the structure of \stdpopsim. The package centers
on a catalog of species (Fig. \ref{fig:cartoon}A), initially consisting of humans, \emph{P.~abelii},
\emph{D.~melanogaster}, \emph{A.~thaliana}, and \emph{E.~coli}.
% RNG: Abbreviate consistently with introduction
A species definition consists of two key elements.  Firstly, the library defines
some basic information about the species' genome, including information about chromosome
lengths, average mutation rates and so on. We also provide access to detailed
empirical information such as genetic maps, which model observed
heterogeneity in recombination rate along chromosomes. As such maps are often large,
we do not distribute them directly with the software, but make them available
for download in a standard format. When a simulation using such a map is
requested by the user, \stdpopsim will transparently download the map
data into a local cache (\texttt{$\sim$/.local/cache/stdpopsim} by default
on Unix platforms), where it can be quickly retrieved for subsequent
simulations.
In the initial version of \stdpopsim we support
the HapMapII~\citep{international2007second} and
deCODE~\citep{kong2010fine} genetic maps in humans;
the \cite{salome2011recombination} map in \emph{A.~thaliana};
and the \cite{comeron2012many} map in \emph{D.~melanogaster}.
The second key element of a species description
within \stdpopsim is a set of carefully curated population genetic model
descriptions from the literature, which are the basis for accurate simulations
of specific historical scenarios that have influenced present-day patterns of
genetic variation. (See the Methods for a description of the community
development and quality-control process for these models.)

Given the genome data and simulation model descriptions defined within the
library, it is then straightforward to run accurate, standardized simulations
across a range of organisms. \stdpopsim has a Python API and a user-friendly
command line interface, allowing users with minimal experience direct access to
state-of-the-art simulations. Simulations are output in the ``tree sequence''
format \citep{kelleher2016efficient,kelleher2018efficient,kelleher2019inferring}, which
contains complete genealogical information about the simulated samples, is
extremely compact, and can be processed efficiently using the \tskit library
\citep{kelleher2016efficient,kelleher2018efficient}. Currently,
\stdpopsim uses the  \texttt{msprime} coalescent simulator \citep{kelleher2016efficient}
as the default simulation engine. We plan to include other simulation
engines to support scenarios that cannot be modelled under the coalescent framework,
and have already implemented \texttt{SLiM} \citep{haller2019tree,haller2019slim} as
an additional backend.

The \stdpopsim command line interface, by default, outputs citation information
for the models, genetic maps and simulation engines used in any particular run.
We hope that this will encourage users to appropriately acknowledge the
resources used in published work, and encourage authors
of simulation models to contribute to our ongoing community-driven development process.
Together with the \stdpopsim version number and the long-term stable identifiers
for population models and genetic maps,
this citation information will result in well-documented and reproducible
simulation workflows. The individual tree sequence files produced by
\stdpopsim also contain complete provenance information including the command
line arguments, operating system environment and versions of key libraries
used.
